---
title: "Customer_Segmentations"
author: "Vivek Singh"
date: "March 19, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}

memory.limit(size=54000)

library(data.table)
library(dplyr)
library(tidyverse)
library(lubridate)
library(feather)
library(zoo)
library(cluster)
library(factoextra)
library(NbClust)
library(gridExtra)

#V4 <-  fread("~\\V23.csv")
  
###subsetting Q1 data

V1 <- select(V4, contains("Q1")
V2 <- select(X, contains("Q1"))
  
```

Using only numeric variables & K-means clustering

```{r K-means Clustering}
gc()
set.seed(23)

# using different cluster values from 2 to 5 for aggregate data

k2 <- kmeans(V1, centers = 2, nstart = 25)
k3 <- kmeans(V1, centers = 3, nstart = 25)
k4 <- kmeans(V1, centers = 4, nstart = 25)
k5 <- kmeans(V1, centers = 5, nstart = 25)

# plots to compare
p1 <- fviz_cluster(k2, geom = "point", data = V1) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = V1) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = V1) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = V1) + ggtitle("k = 5")

library(gridExtra)
grid.arrange(p1, p2, p3, p4, nrow = 2)

gc()

# comupte optimal number of clusters using Elbow method 
wss <- function(k) {
  kmeans(V1, k, nstart = 10 )$tot.withinss
}

# Compute and plot wss for k = 1 to k = 15
k.values <- 1:15

# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")

# compute optimal number of clusters using Gap statistics

#set.seed(123)
#gap_stat <- clusGap(V1, FUN = kmeans, nstart = 25,
#                    K.max = 10, B = 50)


###Based on the grid plot and the Elbow method, using number of clusters = 4

final <- kmeans(V1, 4, nstart = 25)
fviz_cluster(final, data=V1)

```

K-means is not giving distinct clusters. Next adding categorical data and using K-prototype clustering

```{r Add Categorical Data}
#Adding V13 data to Q1 data for additional categorical variables

#V7 <- V1 %>% left_join(V13, by = c("V3"))

# The number of records increased from X to X+K.. Checking for duplicate records

#V13_duplicates <- V13[duplicated(V13$V3),] #K V3 are duplicates
#V1_duplicates <- V1[duplicated(V1$V3),] #No V3 is duplicate

#replacing duplicate V3 with unique value in V13 data

#V13 <- V13[!duplicated(V13$V3),]

#running the join code again without duplicate V3

V7 <- V1 %>% left_join(V13, by = c("V3"))
V7[is.na(V7)]<-0
V7[is.null(V7)]<-0

str(V7)

# converting character variables to factors

character_vars <- lapply(V7, class) == "character"
V7[, character_vars] <- lapply(V7[, character_vars], as.factor)
V7 <- as.data.frame(V7)

rm(V13_duplicates,V1_duplicates)

```
K-prototype clustering using both numeric and categorical data

```{r K-Prototype clustering}

library(clustMixType)

#K-Prototype clustering for Q1 after adding V13 data

wss_kproto <- function(k)
{
  Tots <- c() #initializes variable to accept the results of the loop
  for ( i in k)
  {
    Tots[i] <-  kproto(V7, k=i)$tot.withinss #runs clustering for each of the values of the value ranges passed 
  }
  Tots <- as.data.frame(Tots) #creates data frame of the results
  return(Tots)#returns the result
}

k.values <- 1:10 #value range to pass to the "wss" function

Within_Sum_Sq <-  wss_kproto(k.values) #calling the function, passing the values, and assigning it to object

cluster_WTSS <- cbind(k.values, Within_Sum_Sq) #binding results from the function with the cluster values for plotting 

#plot that shows cluster within sum of squares plotted with number of clusters
cluster_WTSS %>%  ggplot()+
  geom_line(aes(x=k.values, y=Tots))+
  geom_point(aes(x=k.values, y=Tots))+
  labs(x="Number of Clusters", y="Within Sum of Squares")

#run cluster based on best number of clusters from elbow method
clusterResults <- kproto(V7, k=5)
V7$Cluster <- as.factor(clusterResults$cluster)

fwrite(V7, file = "~\\V7.csv", row.names = FALSE)

#show variable distribution grouped by clusters
clprofiles(clusterResults, V7)

distances <- as.data.frame(clusterResults$dists)
clusters <- as.data.frame(clusterResults$cluster)
distance_cluster <- cbind.data.frame(distances, clusters)

#plots observations and colors them by cluster
ggplot(data=distances, aes(x=distances$V1, y=distances$V2, color=clusterResults$cluster))+
geom_point() + xlim(-250, 50000)+ ylim(0,50000)

```
Grouping based on price sensitivity

```{r Price Sensitivity}

V5 <- V7[,c("V3","V9", "Cluster")]


#creating the groups from 1 to 6 based on percentage changes
V5$V10 <- as.factor(case_when(V5$V9<=0 ~"0",
                              V5$V9<5 ~ "1",
                              V5$V9<10 ~"2",
                              V5$V9<15 ~"3",
                              V5$V9<20 ~"4",
                               V5$V9>20 ~"5"))


 V6 <- select(X, contains("Q2"))
 V8 <- V5 %>% left_join(V6, by = c("V3"))
 V8[is.na(V8)] <-0
 
fwrite(V8, file = "~\\V8.csv", row.names = FALSE) 
 
 trial1 <- V8 %>% group_by(Cluster,V10) %>% summarise(n=n()) 
 trial2 <- V8 %>% filter(`V11`>0) %>% 
            group_by(Cluster,V10) %>% summarise(x=n())

 trial1 <- trial1 %>% left_join(trial2, by = c("Cluster","V10"))
 
 trial1[is.na(trial1)] <- 0
 trial1 <- trial1 %>% group_by(Cluster, V10) %>% mutate(V12 = (round(x/n,2)*100))
 
 fwrite(trial1, file = "~\\trial1.csv", row.names = FALSE)
 
```
K-prototypes is givng better clusters.
Validating the trend observed in the clusters using next quarter data

```{r For next quarter}
set.seed(123)

V14 <- select(V4, contains("Q2"))

V21 <- V14 %>% left_join(V13, by = c("V3"))
V21[is.na(V21)]<-0

str(V21)

# converting character variables to factors
character_vars <- lapply(V21, class) == "character"
V21[, character_vars] <- lapply(V21[, character_vars], as.factor)
V21 <- as.data.frame(V21)

#Clustering
Q2_clusterResults <- kproto(V21, k=5)
V21$Cluster <- as.factor(clusterResults$cluster)

#show variable distribution grouped by clusters
clprofiles(clusterResults, V21)

#creating percentage change groups

V15 <- V21[,c("V3","V16", "Cluster")]

V15$V10 <- as.factor(case_when(V15$`V16`<=0 ~"0",
                                             V15$V16<5 ~ "1",
                                             V15$V16<10 ~"2",
                                             V15$V16<15 ~"3",
                                             V15$V16<20 ~"4",
                                             V15$V16>20 ~"5"))


V17 <- select(cl, contains("Q3"))
V18 <- V15 %>% left_join(V17, by = c("V3"))
V18[is.na(V18)] <-0
 
fwrite(V18, file = "~\\V18.csv", row.names = FALSE) 
 
 V19 <- V18 %>% group_by(Cluster,V10) %>% summarise(n=n()) 
 V20 <- V18 %>% filter(`V25`>0) %>% 
            group_by(Cluster,V10) %>% summarise(x=n())

 V19 <- V19 %>% left_join(V20, by = c("Cluster","V10"))
 
 V19[is.na(V19)] <- 0
 V19 <- V19 %>% group_by(Cluster, V10) %>% mutate(V12 = (round(x/n,2)*100))
 
 fwrite(V19, file = "~\\V19.csv", row.names = FALSE)


```